---
title: "R Notebook"
output: html_notebook
---

# Model Selection with JAGS

Load the data set and the necessary libraries and filter the outliers from the dataset, the one with a MEDV value equal to 50.

```{r}
options(warn = - 1)  
library("rjags")
library(Metrics)
library("dplyr")
library(BAS)
housing <- read.csv("housing.csv",header=T)
housing$X=c()
housing <- filter(housing,housing$MEDV!= 50)
summary(housing)
```

Define the design matrix and the outcome vector to use for the JAGS model.

```{r}
# Select a data subset
sub.idx = 1:13  
num.data = nrow(housing)

# Define outcome and design matrix
X <- as.matrix(housing[1:num.data,sub.idx])
Y <- as.vector(housing[1:num.data,14])

# Get dimensions
N <- dim(X)[1]
p <- dim(X)[2]
```

## Spike and Slab

Define the Spike and Slab model over the parameter and consider a normal distribution for the single data with mean depending on the parameters.

```{r}
# Define model
cat(
  "
  model {
    # Likelihood 
    for (i in 1:N) {
    	mu[i] <- beta0 + inprod(X[i,], beta)
    	Y[i] ~ dnorm(mu[i],tau)
    } 
    
    # Tracing the visited model
    for (j in 1:p) {
    	TempIndicator[j] <- g[j]*pow(2, j) 
    }
    mdl <- 1 + sum(TempIndicator[]) # model index in binary coding 
  
    # Gaussian distribution is parametrized in terms of precision parameter
    beta0 ~ dnorm(0, 0.001)
  
    for(j in 1:p) {
    	tprior[j] <- 1 / var_beta[j]
    	bprior[j] <- 0
    }
  
    for(j in 1:p) {
    	beta_temp[j] ~ dnorm(bprior[j], tprior[j])
    	g[j] ~ dbern(theta[j])
    	theta[j] ~ dunif(0,1)
    	beta[j] <- g[j] * beta_temp[j]	
    }
    tau ~ dgamma(0.001,0.001)
  }
  "
, file = "models/SpSl_housing.bug")

# Data to pass to JAGS
data_JAGS_SpSl <- list(N = N, p = p, Y = Y, X = as.matrix(X), var_beta = rep(1, p))

# A list of initial value for the MCMC algorithm 
inits = function() {
  list(beta0 = 0.0, beta_temp = rep(0,p), g = rep(0,p), theta = rep(0.5, p),
       .RNG.seed = 321, .RNG.name = 'base::Wichmann-Hill') 
}

# Compile model (+ adaptation)
model <- jags.model("models/SpSl_housing.bug", data = data_JAGS_SpSl,
                    n.adapt = 1000, inits = inits, n.chains = 1) 
```

```{r}
# if we want to perform a larger burn in with not adaptation.
cat("  Updating...\n")
update(model,n.iter=5000)

# Posterior parameters JAGS has to track
param <- c("beta0", "beta", "g", "mdl")

# Number of iterations & thinning
nit <- 50000
thin <- 10
```

```{r}
# Sampling (this may take a while)
cat("  Sampling...\n")
outputSpSl <- coda.samples(model = model,
                       variable.names = param,
                       n.iter = nit,
                       thin = thin)

# Save the chain
save(output, file = 'chains/SpSl.dat')
```

Traces of the MCMC obtained.

```{r}
# Plot command for coda::mcmc objects (4 at a time for visual purposes)
for(K in 0:13){
  plot(outputSpSl[,(2*K+1):(2*(K+1))])
}
```

```{r}
# Summary command for coda::mcmc objects
summary(outputSpSl)
```

```{r}
# Cast output as matrix
outputSpSl <- as.matrix(outputSpSl)
```

```{r}
param = outputSpSl[,c(14,1:13)]
param_mean = apply(param, 2, "mean")
param_confint =apply(param, 2, quantile, c(0.025, 0.975))
plot(param_mean, main="Spike and Slab",ylim = c(0,50),xlim = c(1,14),cex.axis = 0.7,pch = 16,xaxt = "n")
axis(1, at=1:14, labels=c("Intercept",colnames(X)),cex.axis = 0.5) 
for(i in 1:14){
  arrows(i, param_confint[1,i], i, param_confint[2,i], length=0.05, angle=90, code=3)
}
```

```{r}
# We save the posterior chain of the inclusion variable in post_g
post_g <- as.matrix(outputSpSl[,15:27])
post_beta <- as.matrix(outputSpSl[,1:13])
apply(post_g, 2, "mean")
post_mean_g <- apply(post_g, 2, "mean") 
```

### Median Probability Model

We consider the posterior mean of the inclusion variable for each parameter and keep all parameter that have mean higher than $50\%$

```{r}
library(ggplot2)
# Plot
df <- data.frame(value = post_mean_g, var = colnames(X))
p1 <- ggplot(data = df, aes(y = value, x = var, fill = var)) + 
  geom_bar(stat="identity") + 
  geom_hline(mapping = aes(yintercept = .5), col = 2, lwd = 1.1) +
  coord_flip() + theme_minimal() + theme(legend.position="none") + 
  ylab("Posterior Inclusion Probabilities") + xlab("")
p1
```

```{r}
# Select best model according to MPM
mp_SpSl <- as.vector(which(post_mean_g > 0.5))
post_mean_g[mp_SpSl]
print(colnames(X)[mp_SpSl])
```

### Highest Posterior Density Model

In the HPD model we select the model that was selected with the highest frequency in the Markov Chain.

```{r}
# Plot the mdl chain
plot(outputSpSl[,"mdl"], pch = 20, xlab = "Iteration", ylab = "Model")
```

```{r}
# Number of models visited
length(unique( outputSpSl[,"mdl"]))
```

```{r}
# Post frequency of visited models
visited_models <- sort(table(outputSpSl[,"mdl"]), decreasing = TRUE)
barplot(visited_models, xlab = "N° Model", ylab = "Posterior Frequency")
```

```{r}
# Getting the unique profiles and sort the results
unique_model <- unique(post_g, MARGIN = 1)
freq <- apply(unique_model, 1,
              function(b) sum(apply(post_g, MARGIN = 1, function(a) all(a == b))))
cbind(unique_model[order(freq,decreasing = T),], sort(freq,decreasing = T))
```

```{r}
colnames(X)[as.logical(unique_model[which.max(freq),])]
HDP_SpSl <- c(1:13)[as.logical(unique_model[which.max(freq),])]
```

### Prediction

For the prediction we use jags to fit the data using the subset of parameters that were found with the different criteria.

For the prediction we consider the standard 30% -70% division of the dataset.

#### MPD

```{r}

# Define model
cat(
  "
  model {
    # Likelihood 
    for (i in 148:N) {
    	mu[i] <- beta0 + inprod(X[i,], beta)
    	Y[i] ~ dnorm(mu[i],tau)
    } 
    
    #pred
    for (i in 1:147) {
    	pred_mu[i] <- beta0 + inprod(X[i,], beta)
    	pred_Y[i] ~ dnorm(pred_mu[i],tau)
    } 
    # Gaussian distribution is parametrized in terms of precision parameter
    beta0 ~ dnorm(0, 0.001)
  
    for(j in 1:p) {
    	tprior[j] <- 1 / var_beta[j]
    	bprior[j] <- 0
    }
  
    for(j in 1:p) {
    	beta[j] ~ dnorm(bprior[j], tprior[j])
    }
    tau ~ dgamma(0.001,0.001)
  }
  "
, file = "models/pred_housing.bug")

# Data to pass to JAGS
data_JAGS_pred <- list(N = N, p = length(mp_SpSl), Y = Y, X = as.matrix(X)[,mp_SpSl], var_beta = rep(1, length(mp_SpSl)))

# A list of initial value for the MCMC algorithm 
inits = function() {
  list(beta0 = 0.0,.RNG.seed = 321, .RNG.name = 'base::Wichmann-Hill') 
}

# Compile model (+ adaptation)
model <- jags.model("models/pred_housing.bug", data = data_JAGS_pred,
                    n.adapt = 1000, inits = inits, n.chains = 1) 

```

```{r}
# if we want to perform a larger burn in with not adaptation.
cat("  Updating...\n")
update(model,n.iter=5000)

# Posterior parameters JAGS has to track
param <- c("beta0", "beta", "pred_Y")

# Number of iterations & thinning
nit <- 50000
thin <- 10
```

```{r}
# Sampling (this may take a while)
cat("  Sampling...\n")
outputMPD <- coda.samples(model = model,
                       variable.names = param,
                       n.iter = nit,
                       thin = thin)

# Save the chain
save(output, file = 'chains/pred.dat')
```

```{r}
# Summary command for coda::mcmc objects
summary(output)
```

```{r}
outMPD <- as.matrix(outputMPD)
param = outMPD[,c(length(mp_SpSl)+1,1:length(mp_SpSl))]
param_mean = apply(param, 2, "mean")
param_confint =apply(param, 2, quantile, c(0.025, 0.975))
plot(param_mean, main="MDP",ylim = c(0,50),xlim = c(1,14),cex.axis = 0.7,pch = 16,xaxt = "n")
axis(1, at=1:(length(mp_SpSl)+1), labels=c("Intercept",colnames(X)[mp_SpSl]),cex.axis = 0.5) 
for(i in 1:(length(mp_SpSl)+1)){
  arrows(i, param_confint[1,i], i, param_confint[2,i], length=0.05, angle=90, code=3)
}
```

```{r}
# Cast output as matrix
pred = outMPD[,(length(mp_SpSl)+2):(length(mp_SpSl)+148)]
pred_mean = apply(pred, 2, "mean")
pred_confint =apply(pred, 2, quantile, c(0.025, 0.975))

aa = aa=hist(pred[,1],seq(-100,100,1),probability=TRUE,)
```

```{r}
#BPM <- predict(medv.bestBIC, estimator = "BPM", newdata=newdata,se.fit = TRUE)
#conf.fit <- confint(pred, parm = "mean")
#conf.pred <- confint(pred, parm = "pred")
plot(pred_mean, main="Out of sample: pred. (black) vs true (red)",ylim=c(0,50),cex.axis = 0.7,pch = 16)
for(i in 1:147){
  arrows(i, pred_confint[1,i], i, pred_confint[2,i], length=0.05, angle=90, code=3)
}
points(seq(1:147),Y[1:147],col="red")

rmse = rmse(pred_mean,Y[1:147])
mae = mae(pred_mean,Y[1:147])
print(rmse)
print(mae)
```

#### HPD

```{r}
# Data to pass to JAGS
data_JAGS_pred <- list(N = N, p = length(HDP_SpSl), Y = Y, X = as.matrix(X)[,HDP_SpSl], var_beta = rep(1, length(HDP_SpSl)))

# A list of initial value for the MCMC algorithm 
inits = function() {
  list(beta0 = 0.0,.RNG.seed = 321, .RNG.name = 'base::Wichmann-Hill') 
}

# Compile model (+ adaptation)
model <- jags.model("models/pred_housing.bug", data = data_JAGS_pred,
                    n.adapt = 1000, inits = inits, n.chains = 1) 
```

```{r}
# if we want to perform a larger burn in with not adaptation.
cat("  Updating...\n")
update(model,n.iter=5000)

# Posterior parameters JAGS has to track
param <- c("beta0", "beta", "pred_Y")

# Number of iterations & thinning
nit <- 50000
thin <- 10
```

```{r}
# Sampling (this may take a while)
cat("  Sampling...\n")
outputHPD <- coda.samples(model = model,
                       variable.names = param,
                       n.iter = nit,
                       thin = thin)

# Save the chain
save(output, file = 'chains/pred.dat')
```

```{r}
# Summary command for coda::mcmc objects
summary(output)
```

```{r}
outHPD <- as.matrix(output)
param = outHPD[,c(length(HDP_SpSl)+1,1:length(HDP_SpSl))]
param_mean = apply(param, 2, "mean")
param_confint =apply(param, 2, quantile, c(0.025, 0.975))
plot(param_mean, main="HDP",ylim = c(0,50),xlim = c(1,14),cex.axis = 0.7,pch = 16,xaxt = "n")
axis(1, at=1:(length(HDP_SpSl)+1), labels=c("Intercept",colnames(X)[HDP_SpSl]),cex.axis = 0.5) 
for(i in 1:(length(HDP_SpSl)+1)){
  arrows(i, param_confint[1,i], i, param_confint[2,i], length=0.05, angle=90, code=3)
}
```

```{r}
# Cast output as matrix
pred = outHPD[,(length(HDP_SpSl)+2):(length(HDP_SpSl)+148)]
pred_mean = apply(pred, 2, "mean")
pred_confint =apply(pred, 2, quantile, c(0.025, 0.975))

aa = aa=hist(pred[,1],seq(-100,100,1),probability=TRUE,)
```

```{r}
#BPM <- predict(medv.bestBIC, estimator = "BPM", newdata=newdata,se.fit = TRUE)
#conf.fit <- confint(pred, parm = "mean")
#conf.pred <- confint(pred, parm = "pred")
plot(pred_mean, main="Out of sample: pred. (black) vs true (red)",ylim = c(0,50),cex.axis = 0.7,pch = 16)
for(i in 1:147){
  arrows(i, pred_confint[1,i], i, pred_confint[2,i], length=0.05, angle=90, code=3)
}
points(seq(1:147),Y[1:147],col="red")

rmse = rmse(pred_mean,Y[1:147])
mae = mae(pred_mean,Y[1:147])
print(rmse)
print(mae)
```
